---
title: "biotidy"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{biotidy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

`biotidy` package offers useful utilities for integrating Bioinformatic objects
such as SummarizedExperiment and Seurat into the tidy(verse) framework. 

It is important to note that `biotidy` serves a different purpose compared to
`[tidyomics](https://github.com/tidyomics)`, which primarily utilizes a
pipe-based workflow for managing bioinformatic objects. Specifically, "biotidy"
provides a method for extracting a data frame from bioinformatic objects,
similar to how the "broom::tidy" function operates on statistical model objects.
The inspiration for `biotidy` came from the functionality of the
`scuttle::makePerCellDF` and `scuttle::makePerFeatureDF` functions on the
`SingleCellExperiment` object. The `biotidy` package extends these functions to
work with other bioinformatic objects, such as `SummarizedExperiment`,
`ExpressionSet`, and `Seurat`. 

```{r setup}
library(biotidy)
mocked_se <- mockSE()
mocked_sce <- mockSCE()
mocked_es <- mockES()
mocked_seurat <- mockSeurat()
```

# makePerCellDF
`makePerCellDF` creates a per-cell data.frame (i.e., where each row represents a
sample / cell) from the Bioinformatic objects.

```{r}
# SummarizedExperiment method
makePerCellDF(mocked_se)[1:5, 1:5]
makePerCellDF(mocked_se, melt = TRUE)[1:5, 1:5]
makePerCellDF(mocked_se, melt = TRUE, keep_rownames = TRUE)[1:5, 1:5]
makePerCellDF(mocked_se, features = FALSE, use_coldata = FALSE)
```


```{r}
# SingleCellExperiment method
makePerCellDF(mocked_sce)[1:5, 1:5]
makePerCellDF(mocked_sce, melt = TRUE)[1:5, 1:5]
makePerCellDF(mocked_sce, melt = TRUE, keep_rownames = TRUE)[1:5, 1:5]
makePerCellDF(mocked_sce, features = FALSE, use_coldata = FALSE)
```


```{r}
# ExpressionSet method
makePerCellDF(mocked_es)[1:5, 1:5]
makePerCellDF(mocked_es, melt = TRUE)[1:5, 1:5]
makePerCellDF(mocked_es, melt = TRUE, keep_rownames = TRUE)[1:5, 1:5]
makePerCellDF(mocked_es, features = FALSE, use_coldata = FALSE)
```

```{r}
# Seurat method
makePerCellDF(mocked_seurat, layer = "counts")[1:5, 1:5]
makePerCellDF(mocked_seurat, layer = "counts", melt = TRUE)[1:5, 1:5]
makePerCellDF(mocked_seurat,
  layer = "counts", melt = TRUE, keep_rownames = TRUE
)[1:5, 1:5]
makePerCellDF(mocked_seurat,
  layer = "counts", features = FALSE, use_coldata =
    FALSE
)
```

# makePerFeatureDF
`makePerFeatureDF` Create a per-feature data.frame (i.e., where each row
represents a feature / gene). 

```{r}
# SummarizedExperiment method
makePerFeatureDF(mocked_se)[1:5, 1:5]
makePerFeatureDF(mocked_se, melt = TRUE)[1:5, ]
makePerFeatureDF(mocked_se, melt = TRUE, keep_rownames = TRUE)[1:5, ]
makePerFeatureDF(mocked_se, features = FALSE, use_rowdata = FALSE)
```


```{r}
# SingleCellExperiment method
makePerFeatureDF(mocked_sce)[1:5, 1:5]
makePerFeatureDF(mocked_sce, melt = TRUE)[1:5, ]
makePerFeatureDF(mocked_sce, melt = TRUE, keep_rownames = TRUE)[1:5, ]
makePerFeatureDF(mocked_sce, features = FALSE, use_rowdata = FALSE)
```


```{r}
# ExpressionSet method
makePerFeatureDF(mocked_es)[1:5, 1:5]
makePerFeatureDF(mocked_es, melt = TRUE)[1:5, ]
makePerFeatureDF(mocked_es, melt = TRUE, keep_rownames = TRUE)[1:5, ]
makePerFeatureDF(mocked_es, features = FALSE, use_rowdata = FALSE)
```

```{r}
# Seurat method
makePerFeatureDF(mocked_seurat, layer = "counts")[1:5, 1:5]
makePerFeatureDF(mocked_seurat, layer = "counts", melt = TRUE)[1:5, ]
makePerFeatureDF(mocked_seurat,
  layer = "counts", melt = TRUE, keep_rownames = TRUE
)[1:5, ]
makePerFeatureDF(mocked_seurat,
  layer = "counts", features = FALSE,
  use_rowdata = FALSE
)
```

# Session information {-}

```{r}
sessionInfo()
```
